---
title: "Mobile Telecom Segmentation"
author: ""
output:
  html_document:
    css: AnalyticsStyles/default.css
    theme: paper
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
  pdf_document:
    includes:
      in_header: AnalyticsStyles/default.sty
always_allow_html: yes
---
 
```{r setuplibraries, echo=FALSE, message=FALSE}
suppressWarnings(source("Library/library.R"))
#source("Library/heatmapOutput.R")
#source("Library/MartrixOperations.R")
#source("Library/allPublicLibraries.R")
# Package options
suppressWarnings(ggthemr('fresh'))  # ggplot theme
opts_knit$set(progress=FALSE, verbose=FALSE)
opts_chunk$set(echo=FALSE, fig.align="center", fig.width=10, fig.height=6.35, results="asis")
options(knitr.kable.NA = '')
```
 
# Context of the Project

Customer segmentation is critical to the success of any company, most especially large ones that serve multiple segments. Without adequate segmentation, companies waste resources pushing products to the wrong set of customers. For our project, we are helping a company to properly segment its customer base in order to improve targeted communication, as well as top and bottom-line performance. 

The client is a mobile telecom operator whose core business is in the B2C prepaid business. Thus our focus will be exclusively on customers who buy prepaid packages (Similar to Starhub prepaid packages). 
Presently, the client uses outdated customer segmentation to design service packages and target the relevant customers with appropriate communication. In recent times, the company has observed its segmentation is obsolete as the results of campaigns yield limited, if any, results. Our goal is to redefine the customer segments to better reflect the company’s business reality.
Given that the data belongs to a real telecom operator (name withheld to maintain confidentiality) which works in a highly competitive market, segmentation profiling data is withheld to preserve the anonymity of the company; focus of the segmentation will be consumption (usage) data.
 
# The Business Question

The objective is to create a new segmentation based on data of customer usage of the different services, so the company can better design and target the right customers with the right packages. Our project's aim is to create that segmentation by answering the following question.
 
<b> What are the main customer segments based on historical usage of the different services (voice, data, etc)? </b>
 
 
# Data Loading & Treatment
 
##1)   Treat and Load the Data
 The team received multiple raw data dump files from the client containing usage and profiling data. Data treatment involved:
<ul><b>a.</b>	merging files using common attributes (customer phone number),</ul> 
<ul><b>b.</b>	masking of confidential data that was essential to the segmentation process (e.g. customer phone number)</ul>
<ul><b>c.</b>	removal of other confidential data fields that were not essential </ul>
<ul><b>d.</b>	renaming columns to mask operator identity </ul>
<ul><b>e.</b>	conversion of all fields to numeric format </ul>
<ul><b>f.</b>	slicing file into chunks and selection of a chunk for segmentation (given the size of the original data file) </ul>

The resulting file was used throughout the segmentation process, which is outlined in the coming paragraphs

 
```{r setupdata1E, echo=TRUE, tidy=TRUE}
# Please ENTER the name of the file with the data used. The file should be a .csv with one row per observation (e.g. person) and one column per attribute. Do not add .csv at the end, make sure the data are numeric.
BDASmallTableClean = "Library/BDASmallTableNew.csv"
 
# Please enter the minimum number below which you would like not to print - this makes the readability of the tables easier. Default values are either 10e6 (to print everything) or 0.5. Try both to see the difference.
MIN_VALUE = 0.5
 
# Please enter the maximum number of observations to show in the report and slides.
# DEFAULT is 10. If the number is large the report may be slow.
max_data_report = 10
```
 
```{r}
ProjectData <- read.csv(BDASmallTableClean)
ProjectData <- data.matrix(ProjectData)
ProjectData_INITIAL <- ProjectData
 
```
 
##2) Describe the Data

In this dataset, we have <b>`r nrow(ProjectData)` rows</b> and <b>`r ncol(ProjectData)` columns </b>. The rows indicate the number of customers we are dealing with in this sample and the columns describe those customers' usage pattern. 
<ul>Onnet: Local call to another customer on the same network </ul>
<ul>Offnet: Call to a customer on another local network </ul>
<ul>Intl: Call to another customer in a different country </ul>
<ul>OG (Outgoing): Activity initiated by customer </ul>
<ul>IC (Incoming): Activity received by customer </ul>
<ul>Data: Internet </ul>
<ul>MOU (Minutes of Usage): Measurement of voice usage in minutes </ul>
<ul>MB (megabytes): Measurement of data usage </ul>
<ul>Inroaming: Customer from another country and network who roams on your network </ul>
<ul>Outroaming: Customer from your network who travels to a foreign country and roams on a foreign network </ul>

 
#Factor Analysis & Component selection
 
In this part we identify the underlying factors that best capture the behaviours of the various variables. We use dimensionality reduction techniques to reduce the many columns into a smaller number of factors that captured at least 80% of the variance within the dataset
 
```{r setupfactor, echo=TRUE, tidy=TRUE}
# Please ENTER the original raw attributes to use.
# Please use numbers, not column names, e.g. c(1:5, 7, 8) uses columns 1,2,3,4,5,7,8
factor_attributes_used = c(3:ncol(ProjectData))
 
# Please ENTER the selection criterions for the factors to use.
# Choices: "eigenvalue", "variance", "manual"
factor_selectionciterion = "eigenvalue"
 
# Please ENTER the desired minumum variance explained
# (Only used in case "variance" is the factor selection criterion used).
minimum_variance_explained = 80  # between 1 and 100
 
# Please ENTER the number of factors to use
# (Only used in case "manual" is the factor selection criterion used).
manual_numb_factors_used = 10
 
# Please ENTER the rotation eventually used (e.g. "none", "varimax", "quatimax", "promax", "oblimin", "simplimax", and "cluster" - see help(principal)). Default is "varimax"
rotation_used = "varimax"
 
```
 
```{r}
factor_attributes_used <- intersect(factor_attributes_used, 1:ncol(ProjectData))
ProjectDataFactor <- ProjectData[1:nrow(ProjectData)-1,factor_attributes_used]
ProjectDataFactor <- ProjectData <- data.matrix(ProjectDataFactor)
```

 
## Steps 1-2: Check the Data
 
We are going to start by doing some basic visual exploration of a few data:
 
```{r}
rownames(ProjectDataFactor) <- paste0("Obs.", sprintf("%02i", 1:nrow(ProjectDataFactor)))
iprint.df(t(head(round(ProjectDataFactor, 2), max_data_report)), scale = TRUE)
```
 
The data we use here have the following descriptive statistics:
 
```{r}
iprint.df(round(my_summary(ProjectDataFactor), 2), scale = FALSE)
```
 
## Step 3: Check Correlations
 
This is the correlation matrix of the customer responses to the `r ncol(ProjectDataFactor)` attitude questions - which are the only questions that we will use for the segmentation (see the case):
 
```{r}
thecor = round(cor(ProjectDataFactor),2)
iprint.df(round(thecor,2), scale=TRUE)
```
 
 
## Step 4: Choose number of factors
 
The dataset recorded all possible types of usage that can be made by a customer. The data allows us to understand the underlying behavior of the customer by doing a factor analysis. The factors will identify the behaviors of customers based on their usage pattern. 
 

```{r}
# Principal Function
UnRotated_Results<-principal(ProjectDataFactor, nfactors=ncol(ProjectDataFactor), rotate="none",score=TRUE)
UnRotated_Factors<-round(UnRotated_Results$loadings,2)
UnRotated_Factors<-as.data.frame(unclass(UnRotated_Factors))
colnames(UnRotated_Factors)<-paste("Comp",1:ncol(UnRotated_Factors),sep="")
```
 
```{r}
# PCA Function
Variance_Explained_Table_results<-PCA(ProjectDataFactor, graph=FALSE)
Variance_Explained_Table<-Variance_Explained_Table_results$eig
Variance_Explained_Table_copy<-Variance_Explained_Table
 
rownames(Variance_Explained_Table) <- paste("Component", 1:nrow(Variance_Explained_Table), sep=" ")
colnames(Variance_Explained_Table) <- c("Eigenvalue", "Pct of explained variance", "Cumulative pct of explained variance")
```
 
Let's look at the **variance explained** as well as the **eigenvalues** :
 
```{r}
iprint.df(round(Variance_Explained_Table, 2), scale = FALSE)
```
 
```{r}
eigenvalues  <- Variance_Explained_Table[, "Eigenvalue"]
df           <- cbind(as.data.frame(eigenvalues), c(1:length(eigenvalues)), rep(1, length(eigenvalues)))
colnames(df) <- c("eigenvalues", "components", "abline")
iplot.df(melt(df, id="components"))
```

Based on the outcome of the factor analysis, we chose to move ahead with the 1st 21 factors, which explained more than 80% of the data. 

## Step 5: Interpret the factors
 
Let's now see how the factors look like:
 
```{r}
if (factor_selectionciterion == "eigenvalue")
  factors_selected = sum(Variance_Explained_Table_copy[,1] >= 1)
if (factor_selectionciterion == "variance")
  factors_selected = 1:head(which(Variance_Explained_Table_copy[,"cumulative percentage of variance"]>= minimum_variance_explained),1)
if (factor_selectionciterion == "manual")
  factors_selected = manual_numb_factors_used
```
 
To better visualize them, we will use what is called a "rotation". In this case we selected the `r rotation_used` rotation. For our data, the `r factors_selected` selected factors look as follows after this rotation:
 
```{r}
Rotated_Results<-principal(ProjectDataFactor, nfactors=max(factors_selected), rotate=rotation_used,score=TRUE)
Rotated_Factors<-round(Rotated_Results$loadings,2)
Rotated_Factors<-as.data.frame(unclass(Rotated_Factors))
colnames(Rotated_Factors)<-paste("Comp.",1:ncol(Rotated_Factors),sep="")
 
sorted_rows <- sort(Rotated_Factors[,1], decreasing = TRUE, index.return = TRUE)$ix
Rotated_Factors <- Rotated_Factors[sorted_rows,]
 
iprint.df(Rotated_Factors, scale=TRUE)
```
 
To better visualize and interpret the factors we "suppressed" loadings with small values, e.g. with absolute values smaller than 0.5. In this case our factors look as follows after suppressing the small numbers:
 
```{r}
Rotated_Factors_thres <- Rotated_Factors
Rotated_Factors_thres[abs(Rotated_Factors_thres) < MIN_VALUE]<-NA
colnames(Rotated_Factors_thres)<- colnames(Rotated_Factors)
rownames(Rotated_Factors_thres)<- rownames(Rotated_Factors)
 
iprint.df(Rotated_Factors_thres, scale=TRUE)
```

To understand if there was any redundancy between factors (i.e. high correlation) we took a look at correlation between the rotated factors 
```{r}
factcor = round(cor(Rotated_Factors),2)
iprint.df(round(factcor,2), scale=TRUE)
```
 
## Step 6:  Save factor scores
 
Now we replace all initial variables used in this part with the factors scores, renaming the columns to reflect the new meanings of the factors we derived. The factor scores look like the following for the first customers in our dataset:
 
```{r}
NEW_ProjectData <- round(Rotated_Results$scores[1:nrow(Rotated_Results$scores)-1,1:factors_selected,drop=F],2)
colnames(NEW_ProjectData)<-paste(c("Offnet Voice + SMS ( < 55%)", "Recharge", "Data Packages", "Out Roaming", "OG Intl Voice", "OG Offnet SMS + Intl IC SMS ( < 55%)", "Voice Package", "VAS Package","OG Intl SMS", "SMS Package", "Pay-as-you-go Data", "OG SMS Onnet", "Content Provider VAS Package", "Blackberry Package", "Games", "IVR Shortcodes", "Onnet Voice OG", "IC Intl Voice", "OG Airtime Gifts", "Active Days", "IC Airtime Gifts"),1:ncol(NEW_ProjectData),sep=" ")

iprint.df(t(head(NEW_ProjectData, 10)), scale=TRUE)
```
 
 
<hr>\clearpage
 

#Segmentation - Clustering Analysis & Segment Description

Now that we had decided on what factors to use as basis for segmentation, we proceeded to cluster the observations and derived, finally, 3 segments. 

```{r setupcluster, echo=TRUE, tidy=TRUE}
# Please ENTER then original raw attributes to use for the segmentation (the "segmentation attributes")
# Please use numbers, not column names, e.g. c(1:5, 7, 8) uses columns 1,2,3,4,5,7,8
segmentation_attributes_used = NEW_ProjectData[1:21] #c(10,19,5,12,3) 

# Please ENTER the number of clusters to eventually use for this report
numb_clusters_used = 3 #Use this to "play" with the data until finding a suitable segmentation

# Please enter the method to use for the segmentation:
profile_with = "kmeans" #"hclust" or "kmeans"

# Please ENTER the distance metric eventually used for the clustering in case of hierarchical clustering 
# (e.g. "euclidean", "maximum", "manhattan", "canberra", "binary" or "minkowski" - see help(dist)). 
# DEFAULT is "euclidean"
distance_used = "manhattan"

# Please ENTER the hierarchical clustering method to use (options are:
# "ward", "single", "complete", "average", "mcquitty", "median" or "centroid").
# DEFAULT is "ward"
hclust_method = "ward.D"

# Please ENTER the kmeans clustering method to use (options are:
# "Hartigan-Wong", "Lloyd", "Forgy", "MacQueen").
# DEFAULT is "Lloyd"
kmeans_method = "Lloyd"

```

```{r}
# Same as the initial data
ProjectData <- ProjectData_INITIAL

segmentation_attributes_used <- intersect(segmentation_attributes_used, 1:ncol(ProjectData))
#profile_attributes_used <- intersect(profile_attributes_used, 1:ncol(ProjectData))

ProjectData_segment <- ProjectData[,segmentation_attributes_used]
#ProjectData_profile <- ProjectData[,profile_attributes_used]

#ProjectData_scaled <- apply(ProjectData, 2, function(r) if (sd(r)!=0) (r-mean(r))/sd(r) else 0*r)
```

##Step 1: Decide on pair-wise distance metric

We need to define a distance metric that measures how different people (observations in general) are from each other. This can be an important choice. Here are the differences between the observations using the distance metric we selected:

```{r}
euclidean_pairwise <- as.matrix(dist(head(NEW_ProjectData, max_data_report), method="manhattan"))
euclidean_pairwise <- euclidean_pairwise*lower.tri(euclidean_pairwise) + euclidean_pairwise*diag(euclidean_pairwise) + 10e10*upper.tri(euclidean_pairwise)
euclidean_pairwise[euclidean_pairwise==10e10] <- NA
rownames(euclidean_pairwise) <- colnames(euclidean_pairwise) <- sprintf("Obs.%02d", 1:max_data_report)

iprint.df(round(euclidean_pairwise))
```


##Step 2: Segmentation and Robustness Analysis
 

We need to select the clustering method to use, as well as the number of clusters. It may be useful to see the dendrogram from Hierarchical Clustering of a sample of the factored data, to have a quick idea of how the data may be segmented and how many segments there may be. **Note that there is vector size limit to hclust method, thus we use a sample of, say, a 500 observations**. 
<br>
Here is the dendrogram for our data:

```{r}
Hierarchical_Cluster_distances <- dist(NEW_ProjectData[1:500,], method=distance_used)
Hierarchical_Cluster <- hclust(Hierarchical_Cluster_distances, method=hclust_method)
# Display dendogram
iplot.dendrogram(Hierarchical_Cluster)
# TODO: Draw dendogram with red borders around the 3 clusters
 #rect.hclust(Hierarchical_Cluster, k=numb_clusters_used, border="red") 
``` 
 
We ploted the "distances" traveled before we need to merge any of the lower and smaller in size clusters into larger ones - the heights of the tree branches that link the clusters as we traverse the tree from its leaves to its root. <br>
**Looking at the first 20, we can see that we can have either 4 or 9 clusters. **

```{r}
num <- nrow(NEW_ProjectData[1:500,]) - 1
df1 <- cbind(as.data.frame(Hierarchical_Cluster$height[length(Hierarchical_Cluster$height):1]), c(1:num))
colnames(df1) <- c("distances","index")
iplot.df(melt(head(df1, 20), id="index"), xlab="Number of Components")
``` 
 
Here is the segment membership of the first `r max_data_report` respondents if we use hierarchical clustering:

```{r}
cluster_memberships_hclust <- as.vector(cutree(Hierarchical_Cluster, k=numb_clusters_used)) 
cluster_ids_hclust=unique(cluster_memberships_hclust)

ProjectData_with_hclust_membership <- cbind(1:length(cluster_memberships_hclust),cluster_memberships_hclust)
colnames(ProjectData_with_hclust_membership)<-c("Observation Number","Cluster_Membership")

iprint.df(round(head(ProjectData_with_hclust_membership, max_data_report), 2))
```

while this is the segment membership if we use k-means:

```{r}
kmeans_clusters <- kmeans(NEW_ProjectData,centers= numb_clusters_used, iter.max=2000, algorithm=kmeans_method)

ProjectData_with_kmeans_membership <- cbind(1:length(kmeans_clusters$cluster),kmeans_clusters$cluster)
colnames(ProjectData_with_kmeans_membership)<-c("Observation Number","Cluster_Membership")

iprint.df(round(head(ProjectData_with_kmeans_membership, max_data_report), 2))
```
 
#Part 3: Profile and interpret the above defined segments

Here we show how the *average* usage behaviour in each usage variable *in each segment* compare to the *average usage behaviour of the whole sample* using the ratio of the two.  

<br>
Here are the usage behaviour of the clusters found above and of the total sample for each of the factors:
```{r}
cluster_memberships_kmeans <- kmeans_clusters$cluster 
cluster_ids_kmeans <- unique(cluster_memberships_kmeans)

if (profile_with == "hclust"){
  cluster_memberships <- cluster_memberships_hclust
  cluster_ids <-  cluster_ids_hclust  
}
if (profile_with == "kmeans"){
  cluster_memberships <- cluster_memberships_kmeans
  cluster_ids <-  cluster_ids_kmeans
}


# WE WILL USE THESE IN THE CLASSIFICATION PART LATER
NewData = matrix(cluster_memberships,ncol=1)

population_average = matrix(apply(NEW_ProjectData, 2, median), ncol=1)
colnames(population_average) <- "Population"
Cluster_Profile_mean <- sapply(sort(cluster_ids), function(i) apply(NEW_ProjectData[(cluster_memberships==i), ], 2, median))
if (ncol(NEW_ProjectData) <2)
  Cluster_Profile_mean=t(Cluster_Profile_mean)
colnames(Cluster_Profile_mean) <- paste("Seg.", 1:length(cluster_ids), sep="")
cluster.profile <- cbind (population_average,Cluster_Profile_mean)

iprint.df(round(cluster.profile, 2))
```

<br>
We measure here the ratios of the average for each cluster to the average of the population, minus 1, (e.g. `avg(cluster)` `/` `avg(population)` `-1`) for each segment and variable for each factor.
<br>
The idea is that if in a segment the average usage behaviour is very different (e.g. away from ratio of 1) than the overall average, then that usage variable may indicate something about the segment relative to the total population.(e.g. if all segments are much like the overall population, there may be no segments). 

```{r}
population_average_matrix <- population_average[,"Population",drop=F] %*% matrix(rep(1,ncol(Cluster_Profile_mean)),nrow=1)
Ratio_result <- Cluster_Profile_mean/population_average_matrix
cluster_profile_ratios2 <- (ifelse(population_average_matrix==0, 0, ifelse(abs(Ratio_result)>=1, 1, Ratio_result)))

colnames(cluster_profile_ratios2) <- paste("Seg.", 1:ncol(cluster_profile_ratios), sep="")
rownames(cluster_profile_ratios2) <- colnames(NEW_ProjectData)

## printing the result in a clean-slate table

iprint.df(round(cluster_profile_ratios-1, 2))
```

<br>
There are 3 segments:
<br>
<br>
**Segment Enterprise**
<br>
Strong in several services that enterprise clients use: offnet and international SMS, voice packages, value added services (VAS) packages, IVR short code (shorter number, with less digits), onnet voice and international voice
<br>
<br>
**Segment Short-term client**
<br>
These customers' key trait was their lower number of active days per month, suggesting that they use the client's network as a secondary option. 
<br>
<br>
**Segment Family abroad**
<br>
These customers are focused on international dialling, showing also strong outroaming usage - these customers travel abraod and may have high value. However, their data usage is quite low, reinforcing the sense that they are voice, and not data customers. 

<br>

